---
title: "Uber Spread Analysis"
author: "Joerio Christo Chandra (702302), Priananda Azhar (702299), Syafiq (275402)"
date: "19-07-2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(ggplot2)
library(ggthemes)
library(lubridate)
library(dplyr)
library(tidyr)
library(DT)
library(factoextra)
options(dplyr.summarise.inform = FALSE)
colors = c("orange", "red", "blue", "green", "purple", "black", "pink")
```

### BACKGROUND

As part of our final project for Data Science course (STIS3023), we run several analyses using Uber's data. Our main objectives are:

* drawing plots
* clustering

### OBJECTIVE

* to get a gist of the Uber service spread based on locations of order
* to find out the trend of customers' order locations based on day, month, hour

### SOURCE

For this project, we use the Uber data that contains latitude and longitude. It is downloaded from <https://data.fivethirtyeight.com/>{Five Thirty Eight}.

### DATA PREPARATION

#### Data Reading
In the beningging, we read the dataset first. We are limiting ourselves to using three CSV files for this project, which is the data of Uber orders on months April to June of the year 2014. The data example of the first 5 rows can be seen below.
```{r echo=FALSE}
apr_data <- read.csv("dataset/uber-raw-data-apr14.csv")
may_data <- read.csv("dataset/uber-raw-data-may14.csv")
jun_data <- read.csv("dataset/uber-raw-data-jun14.csv")
uber_data <- rbind(apr_data, may_data, jun_data)
```

```{r}
head(uber_data, 5)
```

#### Preprocessing
Then, in the prepocessing step, we perform the appropriate formatting of Date.Time column. Then, we proceed to create factors like day, month, year and more. After we reformat the Date.Time column, we created each individual column for hour, minute, second, day, month and year.

```{r}
uber_data$Date.Time <- as.POSIXct(uber_data$Date.Time, format = "%m/%d/%Y %H:%M:%S")
uber_data$Time <- format(uber_data$Date.Time, format="%H:%M:%S")
uber_data$Date.Time <- ymd_hms(uber_data$Date.Time)
uber_data$day <- factor(day(uber_data$Date.Time))
uber_data$month <- factor(month(uber_data$Date.Time, label = TRUE))
uber_data$year <- factor(year(uber_data$Date.Time))
uber_data$dayofweek <- factor(wday(uber_data$Date.Time, label = TRUE))
uber_data$hour <- factor(hour(hms(uber_data$Time)))
uber_data$minute <- factor(minute(hms(uber_data$Time)))
uber_data$second <- factor(second(hms(uber_data$Time)))
```

### DATA ANALYSIS
In the Data analysis step, we will ...

#### Trips by Hours in a Day
Plotting the Uber data which is grouped by hours in a day.
The goal is to see how frequent an Uber ride service is ordered per day.
```{r}
hour_data <- uber_data %>%
  group_by(hour) %>%
  dplyr::summarize(Total = n())

ggplot(hour_data, aes(hour, Total)) + 
  geom_bar(stat = "identity", fill = "steelblue", color = "red") +
  ggtitle("Trips Every Hour") +
  theme(legend.position = "none") +
  scale_y_continuous(labels = scales::comma_format())
```

#### Trips by Hour and Month
```{r}
month_hour <- uber_data %>%
  group_by(month, hour) %>%
  dplyr::summarize(Total = n())

ggplot(month_hour, aes(hour, Total, fill = month)) + 
  geom_bar(stat = "identity") +
  ggtitle("Trips by Hour and Month") +
  scale_y_continuous(labels = scales::comma_format())
```


#### Trips During Every Day of the Month
Plotting the Uber data which is grouped by day.

```{r}
day_group <- uber_data %>%
  group_by(day) %>%
  dplyr::summarize(Total = n())

ggplot(day_group, aes(day, Total)) + 
  geom_bar(stat = "identity", fill = "steelblue") +
  ggtitle("Trips Every Day") +
  theme(legend.position = "none") +
  scale_y_continuous(labels = scales::comma_format())
```


```{r}
day_month_group <- uber_data %>%
  group_by(month, day) %>%
  dplyr::summarize(Total = n())
ggplot(day_month_group, aes(day, Total, fill = month)) + 
  geom_bar( stat = "identity") +
  ggtitle("Trips by Day and Month") +
  scale_y_continuous(labels = scales::comma_format()) +
  scale_fill_manual(values = colors)
```

```{r}
month_group <- uber_data %>%
  group_by(month) %>%
  dplyr::summarize(Total = n())
ggplot(month_group, aes(month, Total, fill = month)) + 
  geom_bar( stat = "identity") +
  ggtitle("Trips by Month") +
  theme(legend.position = "none") +
  scale_y_continuous(labels = scales::comma_format()) +
  scale_fill_manual(values = colors)
```

```{r}
month_weekday <- uber_data %>%
  group_by(month, dayofweek) %>%
  dplyr::summarize(Total = n())
ggplot(month_weekday, aes(month, Total, fill = dayofweek)) + 
  geom_bar( stat = "identity", position = "dodge") +
  ggtitle("Trips by Day and Month") +
  scale_y_continuous(labels = scales::comma_format()) +
  scale_fill_manual(values = colors)
```


#### CALCULATING NUMBER OF TRIPS USING BASES

xxx

```{r}
ggplot(uber_data, aes(Base)) + 
  geom_bar(fill = "darkred") +
  scale_y_continuous(labels = scales::comma_format()) +
  ggtitle("Trips by Bases")
```

```{r}
ggplot(uber_data, aes(Base, fill = month)) + 
  geom_bar(position = "dodge") +
  scale_y_continuous(labels = scales::comma_format()) +
  ggtitle("Trips by Month and Bases") +
  scale_fill_manual(values = colors)
```


#### HeatMap

```{r}
month_base <-  uber_data %>%
  group_by(Base, month) %>%
  dplyr::summarize(Total = n())
ggplot(month_base, aes(Base, month, fill = Total)) +
  geom_tile(color = "white") +
  ggtitle("Heatmap by Month and Bases")
```

#### K-MEANS CLUSTERING
The goal of performing K-Means Clustering is to partition or separate the observations into `k` clusters, in which each of the observations belongs to the cluster with the nearest mean. For the this, we have to specify the attributes that we want to include in the cluster. Our team chose Latitude and Longitude attributes for this cluster.

```{r}
clus <- uber_data %>%
  group_by(Lat, Lon) %>%
  dplyr::summarize()
```

The value of `k` is the number of clusters that we want our data points to be divided into. We are using `k = 6` for this project.

```{r}
k2 <- kmeans(clus, centers = 2, nstart = 25)
str(k2)
fviz_cluster(k2, data = clus, main = "Uber Spread by Cluster with 2 Centers")
```

#### GEOPLOT - MAP BASED ON UBER RIDE ORDERS
Lastly, we visualize the rides in New York city by creating a geo-plot that will help us to visualize the rides during April until June 2014.

```{r echo=FALSE}
min_lat <- 40.5774
max_lat <- 40.9176
min_long <- -74.15
max_long <- -73.7004
```

```{r}
ggplot(uber_data, aes(x=Lon, y=Lat)) +
 geom_point(size=1, color = "blue") +
 scale_x_continuous(limits=c(min_long, max_long)) +
 scale_y_continuous(limits=c(min_lat, max_lat)) +
 theme_map() +
 ggtitle("NYC Map Based On Uber Ride Orders During Apr-Jun 2014")
```

### CONCLUSION
